---
title: "R Code Sample"
author: "Yile Chen"
date: "12/13/2021"
output:
  pdf_document: 
    latex_engine: xelatex
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Construction

```{r}
library(tidyverse)
library(zoo)
library(dplyr)
library(skimr)
library(plm)
library(foreign)
library(xtable)
library(AER)
library(stats)
library(sandwich)
```

### 1.a
5331 observations are dropped in the first step and 8686 observations are dropped in the second step.
```{r}
ccm <- read.csv('ccm_10_19.csv')
```

```{r}
ccm$year <- format(as.Date(as.character(ccm$datadate), format = "%Y%m%d"), "%Y") #extract year
ccm_a1 <- ccm[!is.na(ccm$at),] #check NA prices
dim(ccm)[1]-dim(ccm_a1)[1]
```
```{r}
ccm_a2 <- ccm_a1[ccm_a1$fic == "USA",] #check FIC
dim(ccm_a1)[1]-dim(ccm_a2)[1]
```

### 1.b
BVPS here measures the amount of equity belong to preferred shareholders. we need to substract it from the calculation of BE because preferred shareholders are prioritized in claiming assets and earnings than common shareholders.

```{r}
ccm <- ccm_a2 %>% mutate(bvps = case_when(!is.na(ccm_a2$pstkrv) ~ ccm_a2$pstkrv, 
       !is.na(ccm_a2$pstkl) ~ ccm_a2$pstkl, 
       !is.na(ccm_a2$pstk) ~ ccm_a2$pstk,
       TRUE ~ 0)) #calculate BVPS

ccm$BE = ccm$seq + ccm$txditc - ccm$bvps #calculate book value of equity
```


### 1.c
```{r}
ccm <- ccm %>% mutate(lvg = dltt/at) #calculate leverage
```

```{r}
ccm <- ccm %>% distinct(year, LPERMNO, .keep_all = TRUE)
```


### 1.d
Net income, total dividends, and R&D expenses are represented by NI, DVT and XRD in the dataset.

### 2.a
12030 duplicates dropped.
```{r}
crsp <- read.csv('crsp_07_19.csv')
crsp$year <- format(as.Date(as.character(crsp$date), format = "%Y%m%d"), "%Y") #extract year
crsp$month <- format(as.Date(as.character(crsp$date), format = "%Y%m%d"), "%m") #extract month
```

```{r}
crsp_a <- crsp %>% distinct(year, month, PERMNO, .keep_all = TRUE) #select distinct observations
dim(crsp)[1] - dim(crsp_a)[1] #calculate duplicates
```

### 2.b
Inspecting the summary statistics of the price, I find that around 2.77% of prices are negative. It is because that closing prices on those dates are not available. The price thus takes a value of the average of bid/ask prices, and a minus sign is put before the number to distinguish those price from real closing prices. 

For this task, I took absolute value of the alternate price to solve the problem.
```{r}
summary(crsp_a$PRC)
sum(crsp_a$PRC <= 0, na.rm=TRUE)/dim(crsp_a)[1]

crsp <- crsp_a%>%
  mutate(mktcap = abs(SHROUT * ALTPRC),
         mktcap = if_else(mktcap == 0, as.numeric(NA), mktcap),
         RET = as.numeric(RET),
         VOL = as.numeric(VOL))
```

### 2.c
```{r}
crsp$rv = abs(crsp$RET)/crsp$VOL #add a column calculating |ret|/vol for each observation
illiq <- crsp %>% group_by(year, PERMNO) %>% filter(PRC > 0) %>% #group by year and company and filter by price
                  summarize(ILLIQ = mean(rv, na.rm = TRUE)) #aggregate by mean values
head(illiq)
crsp <- merge(crsp, illiq, by = c("year", "PERMNO"), all = TRUE) #merge to add illiq values to the CRSP dataset
```

### 2.d
I used market equity at end of December of the year to represent annualized market equity.
```{r}
me <- crsp %>% group_by(year, PERMNO) %>% filter(month == 12) %>% 
                summarize(ME = mean(mktcap, na.rm = TRUE)/1000) #build the ME variable
crsp <- merge(crsp, me, by = c("year", "PERMNO"), all = TRUE) #merge to add ME to the CRSP dataset
```

```{r}
#prepare a annual data set with selected variables
crsp_short <- crsp %>% select(PERMNO, year, ME, ILLIQ, NAICS) %>% 
                      distinct(year, PERMNO, .keep_all = TRUE) %>% filter(year > 2009)
```


### 3.a & 3.b
For duplicate entries, I aggregate them by summing up their values. I end up with 223 unique industries.
For 3b, I first extract the diagonal, divide it by T005 values then aggregate the results as I did in a.
```{r}
io <- read.csv("IO2012.csv")
```
```{r}
io_a <- data.frame(colnames(io)[2:407], diag(as.matrix(io[, -1])), as.numeric(io[406, 2:407]))
colnames(io_a) = c("code", "diag", "t005") #build a new df of code, diagnol values, and T005 values.
io_a$share = io_a$diag/io_a$t005 #calculate within industry share
io_b <- io_a %>% mutate(code = ifelse(substr(code, 1, 1) == "X", 
                                      substr(code, 2, 5), 
                                      substr(code, 1, 4))) %>% #transform to 4-digits
                                      group_by(code) %>% 
                                      summarize(share = sum(share, na.rm = TRUE)) #aggregate

dim(io_b)[1]
```



### 4.a
Around 21.94% of NAICS codes of the two datasets are different, and the CRSP dataset has less NA values for NAICS code. 
I therefore decided to use the code in the CRSP dataset - labeled as "NAICS4".
```{r}
ccm$PERMNO = ccm$LPERMNO
df <- merge(ccm, crsp_short, by=c("year", "PERMNO"), all = TRUE) #merge CCM and CRSP
```

```{r}
#compare code of the CCM and CRSP
df$naics4 = substr(df$naics,1,4)
df$NAICS4 = substr(df$NAICS,1,4)
sum(df$NAICS4 != df$naics4, na.rm = TRUE)/dim(df)[1]
sum(is.na(df$NAICS4))
sum(is.na(df$naics4))
```

```{r}
df <- left_join(df, io_b, by = c("NAICS4" = "code"), keep = TRUE) #match within industry shares
```

### 4.b
```{r}
df$BM = df$BE/df$ME #create the book-to-market variable
```

```{r}
#transform the merged dataframe into panel format.
df_p <- pdata.frame(df, index = c("PERMNO","year"), drop.index = FALSE)
```

## Analaysis

### 1.
The summary statistics for book equity (BE), Leverage (lvg), market equity (ME), and book-to-market ratio (BM) are shown below.
```{r}
df_1 <- df %>% select(BE, lvg, ME, BM)
summary(df_1)
```

### 2.
The plot for distribution of R&D expenses among 2-digit NAICS industries is shown below. We can see that industries with 2-digit code ‘33’ has the highest R&D expenditure as a group.
```{r}
df_2012 <- df %>% filter(year == 2012)
df_2012$NAICS2 = substr(df_2012$NAICS, 1, 2) #extract 2-digit
df_2012 <- df_2012 %>% group_by(NAICS2) %>% summarise(XRD = sum(xrd, na.rm=TRUE)) #aggregate
barplot(df_2012$XRD ~ df_2012$NAICS2, xlab = "2-digit Industry", ylab = "R&D expenditure")
```

### 3.
### a).
```{r}
##pre-specify the models
linear_model_ni = ni ~ lag(xrd, 1) + lag(xrd, 2) + lag(xrd, 3) + lag(xrd, 4) + lag(ni, 1) + lag(ME/1000, 1) + lag(BM, 1) + lag(lvg, 1) + lag(ILLIQ, 1)
linear_model_dvt = dvt ~ lag(xrd, 1) + lag(xrd, 2) + lag(xrd, 3) + lag(xrd, 4) + lag(ni) + lag(ME/1000) + lag(BM) + lag(lvg) + lag(ILLIQ)
```

```{r}
#OLS for net income
ni_ols_fit   <- plm(linear_model_ni, df_p, model = "pooling", index = c("PERMNO","year"))
summary(ni_ols_fit)
```

### b).
```{r}
#OLS for dividend
dvt_ols_fit <- plm(linear_model_dvt, df_p, model = "pooling", index = c("PERMNO","year"))
summary(dvt_ols_fit)
```

### c). add industry control
```{r}
#pre-specify the models
linear_model_ni_ind = ni ~ lag(xrd, 1) + lag(xrd, 2) + lag(xrd, 3) + lag(xrd, 4) + lag(ni, 1) + lag(ME/1000000, 1) + lag(BM, 1) + lag(lvg, 1) + lag(ILLIQ, 1) + share
linear_model_dvt_ind = dvt ~ lag(xrd, 1) + lag(xrd, 2) + lag(xrd, 3) + lag(xrd, 4) + lag(ni) + lag(ME/1000000) + lag(BM) + lag(lvg) + lag(ILLIQ) + share
```

```{r}
#OLS with industry control for net income
ni_ols_fit_ind <- plm(linear_model_ni_ind, df_p, model = "pooling", index = c("PERMNO","year"))
summary(ni_ols_fit_ind)
```

```{r}
#OLS with industry control for dividend
dvt_ols_fit_ind <- plm(linear_model_dvt_ind, df_p, model = "pooling", index = c("PERMNO","year"))
summary(ni_ols_fit_ind)
```

### d). Fix effect models:

```{r}
#fixed effect model for net income
ni_fe_fit_ind <- plm(linear_model_ni_ind, df_p, model = "within", effect = "individual", index = c("PERMNO","year"))
summary(ni_fe_fit_ind)
```

```{r}
#fixed effect model for dividend
dvt_fe_fit_ind <- plm(linear_model_dvt_ind, df_p, model = "within", effect = "individual", index = c("PERMNO","year"))
summary(dvt_fe_fit_ind)
```